Brief Announcement: An Improved Lower Bound for Dynamic Balanced Graph Partitioning

### 2020-02-15 paper/paper.tex
Restructuring

@@

\begin{proof}
  If no partition of \TAlg's components exist, then it performs no repartition and the cost is $0$.
  Hence, in the following we bound the cost of moving to component-respecting partition.
  We define three types of clusters, $C_1, C_2, C_3$; in cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.

@@

\begin{theorem}
  \TAlg{} is $O(\ell)$-competitive.
\end{theorem}
\begin{proof}
  Fix a completed phase of \TAlg, and consider the state of its counters at the end of it.
  We say that the pair $(u, v)$ is \emph{saturated}, if the counter has value $\alpha$, and \emph{unsaturated} otherwise.

  We partition the input sequence $\sigma$ that arrived during the phase into subsequences $\sigma_I$ (called \emph{internal}) and $\sigma_E$ (called \emph{external}), defined as follows.
  Requests $\sigma_E$ concern pairs of nodes that belong to the same component of \TAlg at the end of the phase, and $\sigma_E$ denotes all other requests issued during this phase.

  %Let $A^I$ be the cost of (extra-cluster) communication incurred in this phase by \TAlg between pairs that belonged to the single component at the end of the phase.
  %Let $A^E$ be the cost of (extra-cluster) communication incurred in this phase between the nodes that belong to different components at the end of this phase.
  Let $ALG(M)$ be the cost of migrations performed by \TAlg in this phase.

  During a single phase, at most $2\cdot \ell$ component merge operations are performed.
  Exceeding this number would mean that a component of size $4$ would emerge, a contradiction.
  Together with Lemma~\ref{lem:1req}, this gives us $ALG(M) \leq 6\cdot\alpha\cdot\ell$.

  Now we bound $ALG(\sigma_I)$.
  Each component of size $3$ contributes at most $(3\cdot \alpha - 1)$ to $A^I$, as at most $2$ of its pairs are saturated and contribute $\alpha$ each, and the unsaturated pair contributes at most $\alpha-1$.
  Each of $\ell$ clusters of \TAlg contain at most one component of size $3$ (other cluster configurations contribute less: $3$ singletons contribute $0$ and a component of size $2$ contributes at most $\alpha$ to $A^I$), and in total, $ALG(\sigma_I) \leq \ell \cdot (3\cdot \alpha-1)$.

@@

  In total, the cost of \TAlg is at most $ALG(\sigma_I) + ALG(\sigma_E) + ALG(M) \leq 9\alpha\cdot \ell + ALG(\sigma_E)$ during this phase.

@@

  Now we lower bound the cost of $\OPT$ by $\OPT(\sigma_I) +\OPT(\sigma_E)+\OPT(M)$, defined with respect to components of \TAlg.
  We split the cost of $\OPT$ into parts coming from serving $\sigma_I$ and $\sigma_E$.
  While serving these requests, $\OPT$ may perform migrations, and we account for them in both parts.

@@

  We have $OPT(M) + OPT(\sigma_I) \geq \alpha$, as the components of \TAlg could not be partitioned without splitting them, and there exist a pair of nodes with at least $\alpha$ requests between them.
  $\OPT$ either served them remotely or performed a migration.

>>>>

In our analysis, we distinguish among three types of clusters, $C_1, C_2, C_3$. In cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.

@@

\begin{proof}
  If no partition of \TAlg's components exist, then it performs no repartition and the cost is $0$.
  Hence, in the following we bound the cost of moving to component-respecting partition.

@@

\begin{theorem}
  \TAlg{} is $O(\ell)$-competitive.
\end{theorem}
\begin{proof}
  Fix a completed phase of \TAlg, and consider the state of its counters at the end of it.
  We consider the incomplete phase later in this proof.
  We say that the pair $(u, v)$ is \emph{saturated}, if the counter has value $\alpha$, and \emph{unsaturated} otherwise.

  We partition the input sequence $\sigma$ that arrived during the phase into subsequences $\sigma_I$ and $\sigma_E$, called \emph{internal} and \emph{external}, respectively.
  The subsequence $\sigma_E$ contains requests to pairs that belong to the same component of \TAlg{} at the end of the phase, and $\sigma_E$ denotes the remaining requests issued during this phase.

  %Let $A^I$ be the cost of (extra-cluster) communication incurred in this phase by \TAlg between pairs that belonged to the single component at the end of the phase.
  %Let $A^E$ be the cost of (extra-cluster) communication incurred in this phase between the nodes that belong to different components at the end of this phase.
  Let $\TAlg(M)$ be the cost of migrations performed by \TAlg in this phase.
  During a single phase, at most $2\cdot \ell$ component merge operations are performed.
  Exceeding this number would mean that a component of size $4$ would emerge, a contradiction.
  Together with Lemma~\ref{lem:1req}, this allows us to bound the cost of migrations, $\TAlg(M) \leq 6\cdot\alpha\cdot\ell$.

  Now we bound $\TAlg(\sigma_I)$.
  Cluster of type $C_3$ contributes at most $(3\cdot \alpha - 1)$ to $\TAlg(\sigma_I)$, as at most $2$ of pairs of the component are saturated and contribute $\alpha$ each, and the remaining unsaturated pair contributes at most $\alpha-1$.
  Other cluster configurations contribute less: $C_1$ contributes $0$ and $C_2$ contributes $\alpha$ to $\TAlg(\sigma_I)$.
  Summing this over all $\ell$ clusters gives us $\TAlg(\sigma_I) \leq \ell \cdot (3\cdot \alpha-1)$.

@@

  In total, the cost of \TAlg is at most $\TAlg(\sigma_I) + \TAlg(\sigma_E) + \TAlg(M) \leq 9\alpha\cdot \ell + \TAlg(\sigma_E)$ during this phase.

@@

  Now we lower bound the cost of $\OPT$ by $\OPT(\sigma_I) +\OPT(\sigma_E)+\OPT(M)$, with $\sigma_I$ and $\sigma_E$ defined with respect to components of \TAlg, and $\OPT(M)$ denoting the cost of migrations performed by $\OPT$.
  %We split the cost of $\OPT$ into parts coming from serving $\sigma_I$ and $\sigma_E$.
  While serving these requests, $\OPT$ may perform migrations, and we account for them in both parts.

@@

  We have $OPT(M) + OPT(\sigma_I) \geq \alpha$, as the components of \TAlg{} could not be partitioned without splitting them, and there exist a pair of nodes with at least $\alpha$ requests between them.
  $\OPT$ either served them remotely or performed a migration.
